---
title: "Intro-to-ML-Regression-and-ML-Classification"
author: "Reinp"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: 
    keep_md: yes
  word_document: default
---

# R Programming: Linear Regression (OLS)

## Set Chunk requirements

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
#echo=FALSE indicates that the code will not be shown in the final document 
#(though any results/output would still be displayed).
#include=FALSE to have the chunk evaluated, but neither the code nor its output displayed
# warning=FALSE and message=FALSE suppress any R warnings or messages from being included 
#in the final document
```

## Load Relevant Packages and Data Set

```{r}
setwd('E:/Documents/Reinp/GitHub Respositories/Intro-to-ML-Regression-and-ML-Classification')

library(stats)
library(psych)
library(tidyverse)
## tidyverse includes readr, ggplot2, dplyr, forcats, tibble, tidyr, purrr, stringr

cars20_tr <- read_csv("data/cars20_train.csv")

view(cars20_tr)

```

## Exploratory Data Analysis

### Exploratory plots in base R 

```{r}

### Histogram of all MPGs

hist(cars20_tr$mpg, breaks = 30)

### Scatter plot of mpg vs disp 

plot(mpg~disp, data = cars20_tr)


### Box plots mpg vs transmission 
boxplot(mpg~transmission, data = cars20_tr)


```

### Exploratory plots in ggplot2

```{r}
### What is the distribution of mpg?

mpg_hist <- ggplot(data = cars20_tr, 
       mapping = aes(
         x = mpg
       ) ) + geom_histogram(fill = "lightblue", colour = "black") +
  theme_bw()

mpg_hist

### What is the relationship between displacement and mpg?

displ_mpg_Scattplt <- ggplot(data = cars20_tr,
       mapping = aes(
         x = disp,
         y = mpg
       )) + geom_point(alpha = 0.3) + 
  geom_smooth()

displ_mpg_Scattplt

### What is the relationship between transmission type and mpg?

trans_mpg_boxplt <- ggplot(data = cars20_tr,
       aes(x = transmission, 
           y = mpg)) + 
  geom_boxplot() + theme_minimal()

trans_mpg_boxplt
```

## Simple Linear Regression

```{r}
## Simple 1 variable model using displacement to predict mpg


model1 <- lm(mpg~disp, data = cars20_tr)

summary(model1)

## To get the model predicted values, use the predict function. 
## It will output a vector of values: one for each car in the training data.

cars20_tr$predictmodel1 <- predict(model1, newdata = cars20_tr)

predict(model1, newdata = cars20_tr, interval = "confidence")

cars20_tr$residualmodel1 <- cars20_tr$predictmodel1 - cars20_tr$mpg


cars20_tr <- cars20_tr %>% mutate(model1 = predict(model1, newdata = cars20_tr),
                                            residual1 = model1 - mpg)

```
### assess the model fit

```{r}

#plot(model1) ### shows several diagnostic graphs

plot(model1, 1)
plot(model1, 2)
plot(model1, 3)  
plot(model1, 4)  
plot(model1, 5)  
plot(model1, 6)

### Showing the fitted model vs the actual mpg values

ggplot(data = cars20_tr, 
       mapping = aes(
        x = mpg,
        y = predictmodel1
       )
) + 
  geom_point() + 
  geom_abline(slope = 1, intercept = 0, colour = "red") + 
  xlim(c(0,60)) + ylim(c(0,60))

### Showing the residual vs the actual mpg values

ggplot(
  data = cars20_tr,
  mapping = aes(
    x = mpg,
    y = residualmodel1
  )
) + 
  geom_point() + 
  geom_hline(yintercept = 0, linetype = "dashed", colour = "red")
```

### Testing on the query set

```{r}
cars20_qu <- read_csv("data/cars20_query.csv")
view(cars20_qu)

cars20_qu$predictmodel1 <- predict(model1, newdata = cars20_qu)
cars20_qu$residualmodel1 <- cars20_qu$predictmodel1 - cars20_qu$mpg


ggplot(data = cars20_qu, 
       mapping = aes(
        x = mpg,
        y = predictmodel1
       )
) + 
  geom_point() + 
  geom_abline(slope = 1, intercept = 0, colour = "blue") + 
  xlim(c(0,60)) + ylim(c(0,60))


ggplot(
  data = cars20_qu,
  mapping = aes(
    x = mpg,
    y = residualmodel1
  )
) + 
  geom_point() + 
  geom_hline(yintercept = 0, linetype = "dashed",  colour = "blue")

```

## Multiple linear regression

```{r}
model2 <- lm(mpg~disp+transmission+drive+atvType, data = cars20_tr) 
model2

summary(model2)

cars20_tr$model2 <- predict(model2, newdata = cars20_tr)

cars20_tr$residual2 <- cars20_tr$model2 - cars20_tr$mpg
```
### assess the model fit

```{r}
plot(model2)


ggplot(data = cars20_tr, 
       mapping = aes(
        x = mpg,
        y = model2
       )
) + 
  geom_point() + 
  geom_abline(slope = 1, intercept = 0, colour = "green") + 
  xlim(c(0,60)) + ylim(c(0,60))


ggplot(
  data = cars20_tr,
  mapping = aes(
    x = mpg,
    y = residual2
  )
) + 
  geom_point() + geom_hline(yintercept = 0, linetype = "dashed", colour = "green")
```

### Testing on the query set

```{r}

cars20_qu$predictmodel2 <- predict(model2, newdata = cars20_qu)
cars20_qu$residualmodel2 <- cars20_qu$predictmodel2 - cars20_qu$mpg


ggplot(data = cars20_qu, 
       mapping = aes(
        x = mpg,
        y = predictmodel2
       )
) + 
  geom_point() + 
  geom_abline(slope = 1, intercept = 0, colour = "yellow") + 
  xlim(c(0,60)) + ylim(c(0,60))


ggplot(
  data = cars20_qu,
  mapping = aes(
    x = mpg,
    y = residualmodel2
  )
) + 
  geom_point() + 
  geom_hline(yintercept = 0, linetype = "dashed",  colour = "yellow")

```




# R Programming: Regression (Decision Trees)

## Making a model with two variables

### Training data

```{r}
library(rpart) ### for decision tree 
library(rpart.plot) ## for showing decision tree

cars20_trdt <- read_csv("data/cars20_train.csv")
view(cars20_trdt)

dt1 <- rpart(mpg~disp+transmission, data = cars20_trdt)
dt1

rpart.plot(dt1)

cars20_trdt$predict1 <- predict(dt1, newdata = cars20_trdt)
cars20_trdt$residual1 <- cars20_trdt$predict1 - cars20_trdt$mpg

## Comparing model values to actual values

ggplot(data = cars20_trdt,
       aes(x = mpg,
           y = predict1)) + 
  geom_point(shape = 21) + 
  xlim(c(0,60)) + 
  ylim(c(0,60)) + 
  geom_abline(slope = 1, intercept = 0, colour = "orange") + theme_bw() + 
  ggtitle("Predict Training set")

## Comparing residual values to actual values

ggplot(data = cars20_trdt,
  aes(x = mpg,
    y = residual1)) + 
  geom_point() + 
  geom_hline(yintercept = 0, linetype = "dashed", colour = "orange") +
  ggtitle("Residual Training set")
```

### Query data

```{r}
cars20_qudt <- read_csv("data/cars20_query.csv")
view(cars20_qudt)

cars20_qudt$predict1 <- predict(dt1, newdata = cars20_qudt)
cars20_qudt$residual1 <- cars20_qudt$predict1 - cars20_qudt$mpg


ggplot(data = cars20_qudt,
       aes(x = mpg,
           y = predict1)) + 
  geom_point(shape = 21) + theme_bw() + 
  xlim(c(0,60)) + ylim(c(0,60)) + 
  geom_abline(slope = 1, intercept = 0) +
  ggtitle("Predict Query set")


ggplot(data = cars20_qudt,
  aes(x = mpg,
    y = residual1)) + 
  geom_point() + 
  geom_hline(yintercept = 0, linetype = "dashed") +
  ggtitle("Residual Query set")
```

## Making a model with all available 

### Training data

```{r}
## create a data set without the make and model variables using dplyr::select

cars20_trdt2 <- read_csv("data/cars20_train.csv") %>%
  select(-make, -model)
view(cars20_trdt2)


dt2 <- rpart(mpg~., data = cars20_trdt2)
rpart.plot(dt2)


cars20_trdt2$predict1 <- predict(dt2, newdata = cars20_trdt2)
cars20_trdt2$residual1 <- cars20_trdt2$predict1 - cars20_trdt2$mpg

ggplot(data = cars20_trdt2,
       aes(x = mpg,
           y = predict1)) + 
  geom_point() + 
  xlim(c(0,60)) + 
  ylim(c(0,60)) + 
  geom_abline(slope = 1, intercept = 0, colour = "purple") + theme_bw() + 
  ggtitle("Predict Training set")



ggplot(data = cars20_trdt2,
  aes(x = mpg,
    y = residual1)) + 
  geom_point() + 
  geom_hline(yintercept = 0, linetype = "dashed", colour = "purple") +
  ggtitle("Residual Training set")

```

### Query data

```{r}

cars20_qudt2 <- read_csv("data/cars20_query.csv")

cars20_qudt2 <- subset(cars20_qudt2, select = -c(make,model)) #Delete columns by name

cars20_qudt2 <- cars20_qudt2[ !(cars20_qudt2$class == "Passenger Van"), ]
view(cars20_qudt2)

### Query data has a factor (class) with new level (Passenger Van) not in train data
#1. delete the Passenger Van rows from query data set and model will run perfectly

#2. You could try updating dt2$xlevels[["class"]] in the model object
# dt2$xlevels[["class"]] <- union(dt2$xlevels[["class"]], levels(cars20_qudt2$class))

#3. exclude (but not remove) "class" from the training data
#dt2 <- rpart(mpg~., data=cars20_trdt2[,!colnames(cars20_trdt2) %in% c("class")])


cars20_qudt2$predict1 <- predict(dt2, newdata = cars20_qudt2)
cars20_qudt2$residual1 <- cars20_qudt2$predict1 - cars20_qudt2$mpg


ggplot(data = cars20_qudt2,
       aes(x = mpg,
           y = predict1)) + 
  geom_point() + 
  xlim(c(0,60)) + 
  ylim(c(0,60)) + 
  geom_abline(slope = 1, intercept = 0, colour = "brown") + theme_bw() + 
  ggtitle("Predict Training set")



ggplot(data = cars20_qudt2,
  aes(x = mpg,
    y = residual1)) + 
  geom_point() + 
  geom_hline(yintercept = 0, linetype = "dashed", colour = "brown") +
  ggtitle("Residual Training set")

```













